### 2018-03-11 me/paper/paper.tex
update

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{times}
\usepackage{titlesec}
\usepackage{geometry}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[colorlinks,linkcolor=blue,filecolor=blue,urlcolor=blue,citecolor=blue]{hyperref}

\renewcommand{\baselinestretch}{1.2}

\title{\large\bfseries Soliton Solution, Ba\"cklund Transformation, Conservation Laws, and rogue wave solution for a generalized variable-coefficient Sasa-Satsuma-like equation}
\author{ZhangShutao}

\begin{document}

\maketitle

\begin{abstract}
A generalized variable-coefficient Sasa-Satsuma-like equation which can describe the 
\end{abstract}

\end{document}