### 2018-11-20 zst/data/reference.tex
update

% !Mode:: "TeX:UTF-8"
\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{参考文献}
\nocite{*}
%\bibliography{data/bibs}

\begin{thebibliography}{0}

\end{thebibliography}
\cleardoublepage