### 2018-11-22 zst/data/reference.tex

% !Mode:: "TeX:UTF-8"
\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{参考文献}
\nocite{*}
%\bibliography{data/bibs}

\begin{thebibliography}{0}

\bibitem{ss-46}
Gilson C , Hietarinta J , Nimmo J , et al. The Sasa-Satsuma higher order nonlinear Schrodinger equation and its bilinearization and multi-soliton solutions[J]. Physical Review E Statistical Nonlinear \& Soft Matter Physics, 2003, 68(2):016614.

\bibitem{ss-45}
Liu Y , Gao Y T , Xu T , et al. Soliton Solution, Bäcklund Transformation, and Conservation Laws for the Sasa-Satsuma Equation in the Optical Fiber Communications[J]. Zeitschrift Für Naturforschung A, 2010, 65(4):291-300.

\bibitem{ss-40}
Chen S . Twisted rogue-wave pairs in the Sasa-Satsuma equation[J]. Phys Rev E Stat Nonlin Soft Matter Phys, 2013, 88(2):023202.

\bibitem{ss-39}
Mu G , Qin Z . Dynamic patterns of high-order rogue waves for Sasa–Satsuma equation[J]. Nonlinear Analysis Real World Applications, 2016, 31:179-209.

\bibitem{nss-1}
Ablowitz M J, Musslimani Z H. Integrable nonlocal nonlinear Schrödinger equation[J]. Physical review letters, 2013, 110(6): 064105.

\bibitem{nss-2}
Yang B, Yang J. Transformations between nonlocal and local integrable equations[J]. Studies in Applied Mathematics, 2018, 140(2): 178-201.

\bibitem{nss-3}
Ma L, Zhao H, Gu H. Integrability and gauge equivalence of the reverse space–time nonlocal Sasa–Satsuma equation[J]. Nonlinear Dynamics, 2018, 91(3): 1909-1920.

\end{thebibliography}
\cleardoublepage