### 2018-11-27 zst/data/reference.tex
update

% !Mode:: "TeX:UTF-8"
\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{参考文献}
\nocite{*}
%\bibliography{data/bibs}

\begin{thebibliography}{0}

\bibitem{syc-1}
武斌. 基于符号计算方法的程序验证技术研究[D]. 华东师范大学, 2010.

\bibitem{syc-3}
Strang G, Strang G, Strang G, et al. Introduction to linear algebra[M]. Wellesley, MA: Wellesley-Cambridge Press, 1993.

\bibitem{syc-5}
张韵华. 符号计算系统 Mathematica 教程[M]. 科学出版社, 2001.

\bibitem{syc-8}
Xu G. Searching for Painlevé integrable conditions of nonlinear PDEs with constant parameters using symbolic computation[J]. Computer Physics Communications, 2008, 178(7): 505-517.

\bibitem{syc-11}
李宏哲. 基于符号计算的 WBK 及其相关方程的达布变换构造和孤子解研究[D]. 北京邮电大学, 2010.

\bibitem{syc-12}
张冬华. 非线性发展方程和孤波解的研究[D]. 北京: 北京邮电大学, 2007.

\bibitem{syc-14}
Liang S, Jeffrey D J. Automatic computation of the travelling wave solutions to nonlinear PDEs[J]. Computer Physics Communications, 2008, 178(9): 700-712.

\bibitem{syc-17}
Zabusky N J, Kruskal M D. Interaction of" solitons" in a collisionless plasma and the recurrence of initial states[J]. Physical review letters, 1965, 15(6): 240.

\bibitem{syc-18}
Gardner C S, Greene J M, Kruskal M D, et al. Method for solving the Korteweg-deVries equation[J]. Physical review letters, 1967, 19(19): 1095.

\bibitem{ss-39}
Mu G, Qin Z. Dynamic patterns of high-order rogue waves for Sasa–Satsuma equation[J]. Nonlinear Analysis: Real World Applications, 2016, 31: 179-209.

\bibitem{ss-40}
Chen S. Twisted rogue-wave pairs in the Sasa-Satsuma equation[J]. Physical Review E, 2013, 88(2): 023202.

\bibitem{ss-41}
Akhmediev N, Soto-Crespo J M, Devine N, et al. Rogue wave spectra of the Sasa–Satsuma equation[J]. Physica D: Nonlinear Phenomena, 2015, 294: 37-42.

\bibitem{ss-46}
Gilson C , Hietarinta J , Nimmo J , et al. The Sasa-Satsuma higher order nonlinear Schrodinger equation and its bilinearization and multi-soliton solutions[J]. Physical Review E Statistical Nonlinear \& Soft Matter Physics, 2003, 68(2):016614.

\bibitem{ss-45}
Liu Y , Gao Y T , Xu T , et al. Soliton Solution, Bäcklund Transformation, and Conservation Laws for the Sasa-Satsuma Equation in the Optical Fiber Communications[J]. Zeitschrift Für Naturforschung A, 2010, 65(4):291-300.

\bibitem{ss-40}
Chen S . Twisted rogue-wave pairs in the Sasa-Satsuma equation[J]. Phys Rev E Stat Nonlin Soft Matter Phys, 2013, 88(2):023202.

\bibitem{ss-39}
Mu G , Qin Z . Dynamic patterns of high-order rogue waves for Sasa–Satsuma equation[J]. Nonlinear Analysis Real World Applications, 2016, 31:179-209.

\bibitem{nss-1}
Ablowitz M J, Musslimani Z H. Integrable nonlocal nonlinear Schrödinger equation[J]. Physical review letters, 2013, 110(6): 064105.

\bibitem{nss-2}
Yang B, Yang J. Transformations between nonlocal and local integrable equations[J]. Studies in Applied Mathematics, 2018, 140(2): 178-201.

\bibitem{nss-3}
Ma L, Zhao H, Gu H. Integrability and gauge equivalence of the reverse space–time nonlocal Sasa–Satsuma equation[J]. Nonlinear Dynamics, 2018, 91(3): 1909-1920.

\end{thebibliography}
\cleardoublepage