### 2018-11-28 zst/data/reference.tex

% !Mode:: "TeX:UTF-8"
\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{参考文献}
\nocite{*}
%\bibliography{data/bibs}

\begin{thebibliography}{0}

\bibitem{syc-1}
武斌. 基于符号计算方法的程序验证技术研究[D]. 华东师范大学, 2010.

\bibitem{syc-3}
Strang G, Strang G, Strang G, et al. Introduction to linear algebra[M]. Wellesley, MA: Wellesley-Cambridge Press, 1993.

\bibitem{syc-5}
张韵华. 符号计算系统 Mathematica 教程[M]. 科学出版社, 2001.

\bibitem{syc-8}
Xu G. Searching for Painlevé integrable conditions of nonlinear PDEs with constant parameters using symbolic computation[J]. Computer Physics Communications, 2008, 178(7): 505-517.

\bibitem{syc-11}
李宏哲. 基于符号计算的 WBK 及其相关方程的达布变换构造和孤子解研究[D]. 北京邮电大学, 2010.

\bibitem{syc-12}
张冬华. 非线性发展方程和孤波解的研究[D]. 北京: 北京邮电大学, 2007.

\bibitem{syc-14}
Liang S, Jeffrey D J. Automatic computation of the travelling wave solutions to nonlinear PDEs[J]. Computer Physics Communications, 2008, 178(9): 700-712.

\bibitem{syc-17}
Zabusky N J, Kruskal M D. Interaction of" solitons" in a collisionless plasma and the recurrence of initial states[J]. Physical review letters, 1965, 15(6): 240.

\bibitem{so-4}
范恩贵. 可积系统与计算机代数[M]. 科学出版社, 2004.

\bibitem{syc-18}
Gardner C S, Greene J M, Kruskal M D, et al. Method for solving the Korteweg-deVries equation[J]. Physical review letters, 1967, 19(19): 1095.

\bibitem{so-1}
広田良吾, 红艳. 孤子理论中的直接方法[M]. 清华大学出版社, 2008.

\bibitem{so-2}
郭玉翠. 非线性偏微分方程引论[M]. 清华大学出版社, 2008.

\bibitem{so-3}
式适, 式达. 物理学中的非线性方程[M]. 北京大学出版社, 2000.

\bibitem{ss-1}
Bandelow U, Akhmediev N. Sasa-Satsuma equation: Soliton on a background and its limiting cases[J]. Physical Review E, 2012, 86(2): 026606.

\bibitem{ss-2}
Chen S. Twisted rogue-wave pairs in the Sasa-Satsuma equation[J]. Physical Review E, 2013, 88(2): 023202.

\bibitem{ss-3}
Ganapathy R, Kuriakose V C, Porsezian K. Soliton propagation in an erbium-doped fiber with and without a continuous wave background[J]. Physical Review E, 2003, 68(6): 066615.

\bibitem{ss-4}
Soto-Crespo J M, Devine N, Hoffmann N P, et al. Rogue waves of the Sasa-Satsuma equation in a chaotic wave field[J]. Physical Review E, 2014, 90(3): 032902.

\bibitem{ss-5}
Zhao L C, Li S C, Ling L. Rational W-shaped solitons on a continuous-wave background in the Sasa-Satsuma equation[J]. Physical Review E, 2014, 89(2): 023210.

\bibitem{ss-6}
Gedalin M, Scott T C, Band Y B. Optical solitary waves in the higher order nonlinear Schrödinger equation[J]. Physical review letters, 1997, 78(3): 448.

\bibitem{ss-7}
Porsezian K, Sundaram P S, Mahalingam A. Coupled higher-order nonlinear Schrödinger equations in nonlinear optics: Painlevé analysis and integrability[J]. Physical Review E, 1994, 50(2): 1543.

\bibitem{ss-8}
Mihalache D, Truta N, Crasovan L C. Painlevé analysis and bright solitary waves of the higher-order nonlinear Schrödinger equation containing third-order dispersion and self-steepening term[J]. Physical Review E, 1997, 56(1): 1064.

\bibitem{ss-9}
Duan L, Zhao L C, Xu W H, et al. Soliton excitations on a continuous-wave background in the modulational instability regime with fourth-order effects[J]. Physical Review E, 2017, 95(4): 042212.

\bibitem{ss-10}
Choudhuri A, Porsezian K. Higher-order nonlinear Schrödinger equation with derivative non-Kerr nonlinear terms: a model for sub-10-fs-pulse propagation[J]. Physical Review A, 2013, 88(3): 033808.

\bibitem{ss-11}
Victor K K, Thomas B B, Kofane T C. Painlevé-integrability of a (2+ 1)-dimensional reaction-diffusion equation: Exact solutions and their interactions[J]. Physical Review E, 2009, 79(5): 056605.

\bibitem{ss-41}
Akhmediev N, Soto-Crespo J M, Devine N, et al. Rogue wave spectra of the Sasa–Satsuma equation[J]. Physica D: Nonlinear Phenomena, 2015, 294: 37-42.

\bibitem{ss-44}
Xu T, Li M, Li L. Anti-dark and Mexican-hat solitons in the Sasa-Satsuma equation on the continuous wave background[J]. EPL (Europhysics Letters), 2015, 109(3): 30006.

\bibitem{ss-46}
Gilson C , Hietarinta J , Nimmo J , et al. The Sasa-Satsuma higher order nonlinear Schrodinger equation and its bilinearization and multi-soliton solutions[J]. Physical Review E Statistical Nonlinear \& Soft Matter Physics, 2003, 68(2):016614.

\bibitem{ss-45}
Liu Y , Gao Y T , Xu T , et al. Soliton Solution, Bäcklund Transformation, and Conservation Laws for the Sasa-Satsuma Equation in the Optical Fiber Communications[J]. Zeitschrift Für Naturforschung A, 2010, 65(4):291-300.

\bibitem{ss-40}
Chen S . Twisted rogue-wave pairs in the Sasa-Satsuma equation[J]. Phys Rev E Stat Nonlin Soft Matter Phys, 2013, 88(2):023202.

\bibitem{ss-39}
Mu G , Qin Z . Dynamic patterns of high-order rogue waves for Sasa–Satsuma equation[J]. Nonlinear Analysis Real World Applications, 2016, 31:179-209.

\bibitem{ss-47}
Liu L, Tian B, Chai H P, et al. Certain bright soliton interactions of the Sasa-Satsuma equation in a monomode optical fiber[J]. Physical Review E, 2017, 95(3): 032202.

\bibitem{ss-48}
Li M, Xu T, Wang L, et al. Nonautonomous solitons and interactions for a variable-coefficient resonant nonlinear Schrödinger equation[J]. Applied Mathematics Letters, 2016, 60: 8-13.

\bibitem{ss-49}
Du Z, Tian B, Wu X Y, et al. Conservation laws and rogue waves for a higher-order nonlinear Schrödinger equation with variable coefficients in the inhomogeneous fiber[J]. Superlattices and Microstructures, 2017, 107: 310-319.

\bibitem{ss-50}
Yuping Z, Junyi W, Guangmei W, et al. The Painlevé property, Bäcklund transformation, Lax pair and new analytic solutions of a generalized variable-coefficient KdV equation from fluids and plasmas[J]. Physica Scripta, 2015, 90(6): 065203.

\bibitem{ss-52}
Weiss J, Tabor M, Carnevale G. The Painlevé property for partial differential equations[J]. Journal of Mathematical Physics, 1983, 24(3): 522-526.

\bibitem{ss-53}
Wazwaz A M, Xu G Q. Modified Kadomtsev–Petviashvili equation in (3+ 1) dimensions: multiple front-wave solutions[J]. Communications in Theoretical Physics, 2015, 63(6): 727.

\bibitem{ss-54}
Ablowitz M J, Kaup D J, Newell A C, et al. Nonlinear-evolution equations of physical significance[J]. Physical Review Letters, 1973, 31(2): 125.

\bibitem{ss-58}
Nakkeeran K, Porsezian K, Sundaram P S, et al. Optical solitons in N-coupled higher order nonlinear Schrödinger equations[J]. Physical review letters, 1998, 80(7): 1425.

\bibitem{ss-60}
Li M , Xu T , Wang L . Dynamical behaviors and soliton solutions of a generalized higher-order nonlinear Schr?dinger equation in optical fibers[J]. Nonlinear Dynamics, 2015, 80(3):1451-1461.

\bibitem{ss-61}
Yin H M, Tian B, Zhen H L, et al. Solitons, bilinear Bäcklund transformations and conservation laws for a-dimensional Bogoyavlenskii-Kadontsev-Petviashili equation in a fluid, plasma or ferromagnetic thin film[J]. Journal of Modern Optics, 2017, 64(7): 725-731.

\bibitem{ss-62}
Ghosh S , Kundu A , Nandy S . Soliton solutions, Liouville integrability and gauge equivalence of Sasa Satsuma equation[J]. Journal of Mathematical Physics, 1999, 40(4):1993.

\bibitem{ss-63}
Wen X Y, Yang Y, Yan Z. Generalized perturbation (n, M)-fold Darboux transformations and multi-rogue-wave structures for the modified self-steepening nonlinear Schrödinger equation[J]. Physical Review E, 2015, 92(1): 012917.

\bibitem{ss-65}
Wen X Y, Yan Z. Higher-order rational solitons and rogue-like wave solutions of the (2+ 1)-dimensional nonlinear fluid mechanics equations[J]. Communications in nonlinear science and numerical simulation, 2017, 43: 311-329.

\bibitem{ss-66}
Wen X Y, Yan Z. Modulational instability and higher-order rogue waves with parameters modulation in a coupled integrable AB system via the generalized Darboux transformation[J]. Chaos: An Interdisciplinary Journal of Nonlinear Science, 2015, 25(12): 123115.

\bibitem{ss-64}
Li M, Liang H, Xu T, et al. Vector rogue waves in the mixed coupled nonlinear Schrödinger equations[J]. The European Physical Journal Plus, 2016, 131(4): 100.

\bibitem{ss-68}
Ma Y C. The perturbed plane‐wave solutions of the cubic Schrödinger equation[J]. Studies in Applied Mathematics, 1979, 60(1): 43-58.

\bibitem{ss-69}
Akhmediev N N, Korneev V I. Modulation instability and periodic solutions of the nonlinear Schrödinger equation[J]. Theoretical and Mathematical Physics, 1986, 69(2): 1089-1093.

\bibitem{nss-1}
Ablowitz M J, Musslimani Z H. Integrable nonlocal nonlinear Schrödinger equation[J]. Physical review letters, 2013, 110(6): 064105.

\bibitem{nss-2}
Yang B, Yang J. Transformations between nonlocal and local integrable equations[J]. Studies in Applied Mathematics, 2018, 140(2): 178-201.

\bibitem{nss-3}
Ma L, Zhao H, Gu H. Integrability and gauge equivalence of the reverse space–time nonlocal Sasa–Satsuma equation[J]. Nonlinear Dynamics, 2018, 91(3): 1909-1920.

\bibitem{nss-4}
Jiang Y, Tian B, Li M, et al. Bright hump solitons for the higher-order nonlinear Schrödinger equation in optical fibers[J]. Nonlinear Dynamics, 2013, 74(4): 1053-1063.

\bibitem{nss-5}
Ma L Y , Zhu Z N . Nonlocal nonlinear Schr\"odinger equation and its discrete version: soliton solutions and gauge equivalence[J]. Journal of Mathematical Physics, 2016, 57(8):064105-171.

\bibitem{nss-6}
Stalin S, Senthilvelan M, Lakshmanan M. Nonstandard bilinearization of PT-invariant nonlocal nonlinear Schrödinger equation: Bright soliton solutions[J]. Physics Letters A, 2017, 381(30): 2380-2385.

\bibitem{nss-7}
Yang B, Yang J. General rogue waves in the nonlocal PT-symmetric nonlinear Schrodinger equation[J]. arXiv preprint arXiv:1711.05930, 2017.

\bibitem{nss-8}
Fokas A S. Integrable multidimensional versions of the nonlocal nonlinear Schrödinger equation[J]. Nonlinearity, 2016, 29(2): 319.

\bibitem{nss-9}
Yang B, Chen Y. Dynamics of high-order solitons in the nonlocal nonlinear Schrödinger equations[J]. Nonlinear Dynamics, 2018: 1-14.

\bibitem{nss-10}
Xu T, Lan S, Li M, et al. Mixed soliton solutions of the defocusing nonlocal nonlinear Schrodinger equation[J]. arXiv preprint arXiv:1804.01021, 2018.

\bibitem{nss-11}
Karpman V I, Rasmussen J J, Shagalov A G. Dynamics of solitons and quasisolitons of the cubic third-order nonlinear Schrödinger equation[J]. Physical Review E, 2001, 64(2): 026614.

\end{thebibliography}
\cleardoublepage